---
title: "Project 1"
author: "Chase LeCroy"
date: "January 12, 2017"
output: html_document
keep_md: true
---
```{r}
require(dplyr)
require(ggplot2)
require(lubridate)
require(lattice)
require(knitr)
require(markdown)
```

####Loading and preprocessing the data  
```{r}
projdata <- read.csv("activity.csv")
projdata <- as.tbl(projdata)
```
No preprocessing was needed.  

####What is mean total number of steps taken per day?  
```{r}
tot_steps_day <- projdata %>% group_by(date) %>% summarize(TotSteps = sum(steps, na.rm=TRUE))
```
```{r fig.height=4}
ggplot(data=tot_steps_day, aes(tot_steps_day$TotSteps)) +labs(x="Total Steps per Day", y="Count")+ geom_histogram()

```

``` {r meanstepsday}
meantot <- mean(tot_steps_day$TotSteps)
mediantot <- median(tot_steps_day$TotSteps)
```

The mean steps per day is `r meantot`.  
The median steps per day is `r mediantot`.  
 

####What is the average daily activity pattern?  
```{r}
activity <- projdata %>% group_by(interval) %>% summarize(MeanSteps = mean(steps, na.rm=TRUE))

plot2 <- ggplot(data=activity, aes(y=activity$MeanSteps, x=activity$interval)) + labs(x="Five Minute Interval", y="Average Steps per Interval")+ geom_line()

plot2

max_steps <- activity[which.max(activity$MeanSteps),1]
```
  
The interval `r max_steps` contains the maximum number of steps on average.  

####Imputing missing values  
```{r}
missing <- sum(is.na(projdata$steps))

projdata_full <- projdata %>% filter(!is.na(steps))
projdata_m <- projdata %>% filter(is.na(steps))
projdata_imp <- left_join(projdata_m, activity, by="interval") %>% select(date,interval, steps=MeanSteps)

projdata_final <- rbind(projdata_full, projdata_imp)

projdata_final_gg <- projdata_final %>% group_by(date) %>% summarize(TotSteps = sum(steps))

plot3 <- ggplot(data=projdata_final_gg, aes(projdata_final_gg$TotSteps)) +labs(x="Total Steps per Day", y="Count")+ geom_histogram()
plot3

meanimp <- mean(projdata_final_gg$TotSteps)
medianimp <- median(projdata_final_gg$TotSteps)
```

There are `r missing` rows with missing values.  
The mean steps per day in the imputed data is `r meanimp`.  
The median steps per day in the imputed data is `r medianimp`.  
Imputing the average steps at each interval for missing data increased both the mean and the median.  

####Are there differences in activity patterns between weekdays and weekends?  

``` {r}
daysdata <- weekdays(parse_date_time(projdata_final$date, "Y-m-d"))
projdata_days <- cbind(projdata_final,daysdata)
projdata_days <- projdata_days %>% mutate(weekday = ifelse(projdata_days$daysdata %in% c("Saturday", "Sunday"), "weekend", "weekday"))

projdata_days_gg <- projdata_days %>% group_by(interval, weekday) %>% summarize(MeanSteps = mean(steps)) 

plot4 <- xyplot(data=projdata_days_gg, MeanSteps~interval|weekday, type ='l', auto.key = T,layout=c(1,2), xlab="Interval",ylab="Mean Steps per Interval")
plot4
```
